```{r include=FALSE}
library(dplyr, quietly = TRUE)
library(ggplot2)
library(paletteer)
```


# Small Worlds and Large Worlds {#worlds}


## 2E1 {-}

$Pr(rain \mid Monday)$ and $\frac{Pr(rain, Monday)}{Pr(Monday)}$

## 2E2 {-}

Prob that it is Monday, given it is raining

## 2E3 {-}

$Pr(Monday \mid rain)$ and $\frac{Pr(rain \mid Monday)Pr(Monday)}{Pr(rain)}$

## 2E4 {-}

Based on what we observe small world, the earth, we conjecture that water
occupies 70% of the space.

## 2M1 {-}

Creating the function to compute the grid

```{r}
# function to calculate the posterior grid as per 2M1
calc_post_water <- function(obs, grid_size, label) {
    
    # Parameters:
    # obs: Character vector of observation with W and L
    # grid_size: Number of grid elements
    # label: String to identify the data, used later to plot

    # convert to boolean values
    obs <- obs == "W"
    
    # create a grid of possible values for the parameter p
    param_grid <- seq(from = 0, to = 1, length.out = grid_size)
    
    # the prior is uniformely distributed
    prior_grid <- dunif(x = param_grid, min = 0, max = 1)
    
    # the likelihood is binomial
    like_grid <- dbinom(x = sum(obs), 
                        size = length(obs), 
                        prob = param_grid)
    
    # compute the posterior
    post_grid <-  like_grid * prior_grid
    # standardize the posterior by dividing by the sum of its possible values
    post_grid <- post_grid / sum(post_grid)
    # add param grid and build a dataframe
    post_grid <- data.frame(param = param_grid, 
                            post = post_grid,
                            label = rep(label, times = grid_size),
                            stringsAsFactors = FALSE)
    
    return(post_grid)
        
}
```


### 2M1 (1) {-}

Compute the grid for the given data using the function defined just abov

```{r}
post_grid_1 <- calc_post_water(c("W", "W", "W"), 10, "(1)")
post_grid_1
```

### 2M1 (2) {-}


```{r}
post_grid_2 <- calc_post_water(c("W", "W", "W", "L"), 10, "(2)")
```

### 2M1 (3) {-}

```{r}
post_grid_3 <- calc_post_water(c("L", "W", "W", "L", "W", "W", "W"), 10, "(3)")
```

### plot {-}

And visualize the results

```{r}
pd <- rbind(post_grid_1, post_grid_2, post_grid_3)  # data for graph
p2M1 <- ggplot(data = pd, mapping = aes(x = param, y = post, color = label)) +
    geom_point(size = 3) + geom_line() +
    scale_color_paletteer_d(palette = "futurevisions::atomic_orange") +
    theme_classic() +
    theme(plot.title = element_text(color = "dark blue", hjust = 0.5),
          legend.position = c(0.1, 0.85)) +
    labs(title = "Exercise 2M1", x = "parameter p", y = "posterior prob")
p2M1
```

## 2M2 {-}

See R code 2.5 in for example of such a prior.

```{r}
# function to calculate the posterior grid as per 2M1
calc_post_water <- function(obs, grid_size, label) {
    
    # Parameters:
    # obs: Character vector of observation with W and L
    # grid_size: Number of grid elements
    # label: String to identify the data, used later to plot
    
    # convert to boolean values
    obs <- obs == "W"
    
    # create a grid of possible values for the parameter p
    param_grid <- seq(from = 0, to = 1, length.out = grid_size)
    
    # the prior is uniformely distributed
    prior_grid <- ifelse(param_grid >= 0.5, 0.1, 0)
    
    # the likelihood is binomial
    like_grid <- dbinom(x = sum(obs), 
                        size = length(obs), 
                        prob = param_grid)
    
    # compute the posterior
    post_grid <-  like_grid * prior_grid
    # standardize the posterior by dividing by the sum of its possible values
    post_grid <- post_grid / sum(post_grid)
    # add param grid and build a dataframe
    post_grid <- data.frame(param = param_grid, 
                            post = post_grid,
                            label = rep(label, times = grid_size),
                            stringsAsFactors = FALSE)
    
    return(post_grid)
}
```

### 2M2 (1) {-}

Compute the grid for the given data using the function defined just abov

```{r}
post_grid_1 <- calc_post_water(c("W", "W", "W"), 10, "(1)")
post_grid_1
```

### 2M2 (2) {-}


```{r}
post_grid_2 <- calc_post_water(c("W", "W", "W", "L"), 10, "(2)")
```

### 2M2 (3) {-}

```{r}
post_grid_3 <- calc_post_water(c("L", "W", "W", "L", "W", "W", "W"), 10, "(3)")
```

### 2M2 plot {-}

And visualize the results

```{r}
pd <- rbind(post_grid_1, post_grid_2, post_grid_3)  # data for graph
p2M2 <- ggplot(data = pd, mapping = aes(x = param, y = post, color = label)) +
    geom_point(size = 3) + geom_line() +
    scale_color_paletteer_d(palette = "futurevisions::atomic_red") +
    theme_classic() +
    theme(plot.title = element_text(color = "dark blue", hjust = 0.5),
          legend.position = c(0.1, 0.85)) +
    labs(title = "Exercise 2M2", x = "parameter p", y = "posterior prob")
p2M2
```

## 2M3 {-}

We first need to formulate the probabilites

$$
\begin{equation}
P(Earth \mid Land) = \\
\frac{P(Earth, Land)}{P(Land)} = \\
\frac{P(Land \mid Earth) \cdot P(Earth)}{P(Land)} = \\
\frac{P(Land \mid Earth) \cdot P(Earth)}
 {P(Land \mid Earth) \cdot P(Earth) + P(Land \mid Mars) \cdot P(Mars)}
\end{equation}
$$

```{r}
prob_earth <- 0.5
prob_mars <- 1 - prob_earth
prob_water_earth <- 0.7
prob_water_mars <- 0
prob_land_earth <- 1 - prob_water_earth
prob_land_mars <- 1 - prob_water_mars

# Pr(Earth | Land) = Pr(Earth, Land) / Pr(Land) = Pr(Land | Earth) * P(Earth) / Pr(Land) =
# Pr(Land | Earth) * P(Earth) / (Pr(Land | Earth) * Pr(Earth) + Pr(Land | Mars) * Pr(Mars))
(prob_land_earth * prob_earth) / 
    (prob_land_earth * prob_earth + prob_land_mars * prob_mars)
```

## 2M4 {-}

We have the random variables

$F$: Represent the face that we see.

* $F=b$: We see the black side.
* $F=w$: We see the white face.

$H$: Represent the face that we *don't see*, the hidden face.

* $H=b$: The hidden side is black.
* $H=w$: The hidden side is white.

The probabilities of $F$ are $P(F=b) = P(F=w) = 0.5$

$C=i$: Represents the card # i that we pick with the following values

* $C=1$: we pick card 1 which is black & black, notation

* $C=2$: we pick card 2 which is black & white

* $C=3$: we pick card 3 which is white & white

and the pdf of $C$ is

|Card|Probability|
|:-------:|:-----------:|
|$C=1$    |$\frac{1}{3}$|
|$C=2$    |$\frac{1}{3}$|
|$C=3$    |$\frac{1}{3}$|


$$
P(H=b \mid F=b) = \frac{P(H=b,F=b)}{P(F=b)} = \frac{P(C=1)}{P(F=b)}= \\
\frac{P(C=1)}{P(F=b \mid C=1)P(C=1) + P(F=b \mid C=2)P(C=2) + P(F=b \mid C=3)P(C=3)} \\
\frac{\frac{1}{3}}{1 \cdot \frac{1}{3} + 0.5 \cdot \frac{1}{3} + 0 \cdot \frac{1}{3}} = \frac{2}{3} \approx 0.67
$$

## 2M5 {-}

Using the same convention as in 2M4 and given that

|Card|Probability|
|:-------:|:-----------:|
|$C=1$    |$\frac{2}{4}$|
|$C=2$    |$\frac{1}{4}$|
|$C=3$    |$\frac{1}{4}$|

then, same as in 2M4 but with different probabilities,

$$
\frac{\frac{2}{4}}{1 \cdot \frac{2}{4} + 0.5 \cdot \frac{1}{4} + 0 \cdot \frac{1}{4}} = \frac{4}{5}=0.8
$$



## 2M6 {-}


|Card|Probability|
|:-------:|:-----------:|
|$C=1$    |$\frac{1}{6}$|
|$C=2$    |$\frac{2}{6}$|
|$C=3$    |$\frac{3}{6}$|


then, same as in 2M4 but with different probabilities,

$$
\frac{\frac{1}{6}}{1 \cdot \frac{1}{6} + 0.5 \cdot \frac{2}{6} + 0 \cdot \frac{3}{6}} = \frac{1}{2}
$$


## 2M7 {-}

$C_i$: The card chosen in the $i$ pick
$F_i$: The face on the $i$ pick
$H_i$: The hidden face of the card on the $i$ pich

The joint probability table is

|$C_1$|$C_2$|$F_1$|$H_1$|$F_2$|$H_2$|freq|prob|
|:---:|:---:|:---:|:---:|:---:|:---:|:--:|:--:
|1|2|b|b|b|w|2|$\frac{2}{24}$|
|1|2|b|b|w|b|2|$\frac{2}{24}$|
|1|3|b|b|w|w|4|$\frac{4}{24}$|
|2|1|b|w|b|b|2|$\frac{2}{24}$|
|2|1|w|b|b|b|2|$\frac{2}{24}$|
|2|3|b|w|w|w|2|$\frac{2}{24}$|
|2|3|w|b|w|w|2|$\frac{2}{24}$|
|3|1|w|w|b|b|4|$\frac{4}{24}$|
|3|2|w|w|b|w|2|$\frac{2}{24}$|
|3|2|w|w|w|b|2|$\frac{2}{24}$|

and using probabilities we have

$$
P(H_1 = b \mid F_1 = b, F_2 = w) = \\
\frac{P(H_1 = b, F_1 = b, F_2 = w)}{P(F_1=b, F_2=w)} = \\
\frac{P(F_2 = w, C_1 = 1)}{P(F_1=b,F_2=w,H_1=b)+P(F_1=b,F_2=w,H_1=w)} =\\
\frac{P(F_2 = w, C_1 = 1)}
 {P(F_2=w,C_1=1)+P(F_1=b,F_2=w,H_1=w)} =\\
\frac{\frac{2}{24} +\frac{4}{24}}
{\frac{2}{24} + \frac{4}{24} + \frac{2}{24}}=\frac{3}{4}=0.75
$$


## 2H1 {-}

We use $A,B$ to identify the species.

We use $T_i=1,0$ to identify if birth at time $i$ is twin $T_i=1$ or singleton $T_i=0$

we have

$$
P(T_i=1 \mid A)=0.1 \\
P(T_i=1 \mid B)=0.2
$$

We need to find $P(B_2=t \mid B_1=t)$ therefore

$$
P(T_2=1 \mid T_1=1)=\\
\frac{P(T_2=1, T_1=1)}{P(T_1=1)}=\\
\frac{P(T_2=1, T_1=1 \mid A) P(A) + P(T_2=1, T_1=1 \mid B) P(B)}{P(T_1=1 \mid A)P(A)+P(T_1=1 \mid B)P(B)}=\\
\frac{0.1 \cdot 0.1 \cdot 0.5 + 0.2 \cdot 0.2 \cdot 0.5}{0.1 \cdot 0.5 + 0.2 \cdot 0.5}=\frac{0.025}{0.15}=\frac{1}{6} \approx 0.17

$$

## 2H2 {-}

$$
P(A \mid T_1 = 1) = \frac{P(A, T_1=1)}{P(T_1=1)} = \\
\frac{P(T_1=1 \mid A)P(A)}{P(T_1=1 \mid A)P(A) + P(T_1=1 \mid B)P(B)} = \\
\frac{0.1 \cdot 0.5}{0.1 \cdot 0.5 + 0.2 \cdot 0.5} = \frac{1}{3} \approx 0.33
$$


## 2H3 {-}

since $T_i$ are independent then

$$
P(A \mid T_1 = 1, T_2 =0)=\\
\frac{P(A,T_1=1,T_2=0)}{P(T_1=1,T_2=0)}=\\
\frac{P(T_1=1, T_2=0 \mid A)P(A)}{P(T_1=1,T_2=0\mid A)P(A) + P(T_1=1,T_2=0\mid B)P(B)}=\\
\frac{0.1 \times 0.9 \times 0.5}{(0.1 \times 0.9 \times 0.5) + (0.2 \times 0.8 \times 0.5)} =
\frac{0.045}{0.045+0.08}=\frac{9}{25}=0.36
$$

## 2H4 {-}

The new genetic test, called $G=a,b$ when it identifies species A or B with a and b.

$$
P(G=a \mid A) = 0.8 \\
P(G=b \mid B) = 0.65 \\
P(G=a \mid B) = 1- P(G=b \mid B) = 1-0.65=0.35
$$

Using the existing prior on the population of bear $P(S=a)=(S=b)=0.5$


$$
P(A | G=a)=\\ 
\frac{P(G=a \mid A) P(A)}{P(G=a)}=\\
\frac{P(G=a \mid A) P(A)}{P(G=a \mid A)P(A) + P(G=a \mid B)P(B)} = \\
\frac{0.8 \cdot 0.5}{0.8 \cdot 0.5 + (1-0.65) \cdot 0.5} = \frac{0.45}{0.45+0.325}=\frac{18}{25} = 0.72
$$


Which means that our prior of $P(A)=0.5$ has been changed to $0.72$ once we use
the genetic test and therefore a better identification.


Similarly we will use the prior without genetic test in exercise 2H3 just above
which had found $P(A') = P(A \mid T_1 = 1, T_2 =0)=0.36$ and use the genetic test as a
new prior $P(A' \mid G=a)$

$$
P(A' \mid G=a) = \\ 
\frac{P(G=a \mid A')P(A')}{P(G=a \mid A')P(A') + P(G=a \mid B')P(B')} \\
\frac{0.8 \cdot 0.36}{0.8 \cdot 0.36 + 0.35 \cdot (1-0.36)} \approx 0.56
$$



