```{r include=FALSE}
library(rethinking)
library(brms)
library(dplyr)
library(tidyr)
library(ggdist)
library(paletteer)
```

# Sampling the Imaginary {#sampling}

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 1000)
prior <- rep(1, length(p_grid))
likelihood <- dbinom(x = 6, size = 9, prob = p_grid)
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample(x = p_grid, size = 1e4, prob = posterior, replace = TRUE)
```


## 3E1 {-#prac3E1}

note: published solution on internet has 5 instead of 4!
I verified the code several times.

The nb of samples below 0.2 is

```{r}
# nb of samples < 0.2
nb <- sum(samples < 0.2)
# the proportion
prop <- nb / length(samples)
sprintf("%d samples representing %0.2f%%", nb, 100 * prop)
```

## 3E2 {-#prac3E2}

```{r}
# nb of samples < 0.8
nb <- sum(samples > 0.8)
# the proportion
prop <- nb / length(samples)
sprintf("%d samples representing %0.2f%%", nb, 100 * prop)
```
## 3E3 {-#prac3E3}

```{r}
# nb of samples between 0.2 ad 0.8
nb <- sum(0.2 <= samples & samples <= 0.8)
# the proportion
prop <- nb / length(samples)
sprintf("%d samples representing %0.2f%%", nb, 100 * prop)
```
## 3E4 {-#prac3E4}

```{r}
quantile(x = samples, probs = 0.2)
```

## 3E5 {-#prac3E5}

```{r}
quantile(x = samples, probs = 1 - 0.2)
```


## 3E6 {-#prac3E6}

This is the posterior high density interval.

with `rethinking`

```{r}
# with rethinking
rethinking::HPDI(samples, prob = 0.66)
```
 and with `ggdist`, we get the mean as a bonus and in a dataframe format.

```{r}
# with ggdist
ggdist::mean_hdi(.data = samples, .width = 0.66)
```

## 3E7 {-#prac3E7}


This is the usual quantile, symetric, interval.

with `rethinking``

```{r}
# with rethinking
rethinking::PI(samples, prob = 0.66)
```

with `ggdist`

```{r}
# with ggdist
ggdist::mean_qi(.data = samples, .width = 0.66)
```


## 3M1 {-#prac3M1}

Create the grid

```{r}
n_success <- 8
n_trials <- 15
grid_size <- 1000
df <- data.frame(
    p_grid = seq(from = 0, to = 1, length.out = grid_size),
    prior = 1) %>%
    mutate(
     likelihood = dbinom(x = n_success, size = n_trials, prob = p_grid),
     posterior = (likelihood * prior) / sum(likelihood * prior)
    )
stopifnot(near(sum(df$posterior), 1))
```


```{r}
ggplot(data = df, mapping = aes(x = p_grid, y = posterior)) +
    geom_point(size = 1, color = "palevioletred", alpha = 0.9) +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(title = "Practice 3M1",
         subtitle = sprintf("Grid size = %d", nrow(df)),
         x = "grid values of probability of water", 
         y = "posterior probability simulated")
```


## 3M2 {-#prac3M2}

Create the sampled data


```{r}
n_samples <- 1e4
set.seed(100)
the_samples <- df %>%
    slice_sample(n = n_samples, weight_by = posterior, replace = TRUE) %>%
    mutate(id = seq_len(n_samples), .before = p_grid)
```

and find the HPDI

```{r}
# HPDI with rethinking
rethinking::HPDI(the_samples$p_grid, prob = 0.9)
```

```{r}
# HPDI with ggdist
ggdist::mean_hdi(.data = the_samples$p_grid, .width = 0.9)
```


## 3M3 {-#prac3M3}


Constructing a posterior prediction by simulating $W$ with the sampled $p$.
See the R code 3.26, p. 66.

```{r}
# using the sampled p we sim
post_pred <- rbinom(n = nrow(the_samples), size = n_trials, prob = the_samples$p_grid)
the_samples <- data.frame(
  the_samples,
  predict = post_pred
)
```

and using it to calculate the prob of having 8 $W$ in 15 trials

```{r}
sum(post_pred == 8 ) / length(post_pred)
```
the histogram of the predictions is

```{r}
colr <- paletteer::paletteer_d("Manu::Hoiho")
ggplot(the_samples, aes(x=predict)) +
  geom_histogram(aes(fill = ..count..)) +
  scale_x_continuous(breaks = 0:15) +
  scale_fill_gradientn(colors = colr) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Practice 3M3",
       subtitle = sprintf("Frequencies of %d predictions", nrow(the_samples)))
```




## 3M4 {-#prac3M4}

Prob of 6 $W$ in 9 tosses with the post dist of 8 $W$ in 15 tosses. Same process as
in 3M3.

```{r}
post_pred <- rbinom(n = nrow(the_samples), size = 9, prob = the_samples$p_grid)
sum(post_pred == 6 ) / length(post_pred)
```

## 3M5 {-#prac3M5}

Create the grid

```{r}
n_success <- 8
n_trials <- 15
grid_size <- 1000
df <- data.frame(
    p_grid = seq(from = 0, to = 1, length.out = grid_size),
    prior = ifelse(p_grid < 0.5, 0, 1)) %>%
    mutate(
     likelihood = dbinom(x = n_success, size = n_trials, prob = p_grid),
     posterior = (likelihood * prior) / sum(likelihood * prior)
    )
stopifnot(near(sum(df$posterior), 1))
```


```{r}
ggplot(data = df, mapping = aes(x = p_grid, y = posterior)) +
  geom_point(size = 1, color = "slateblue1", alpha = 0.9) +
  geom_line(color = "slateblue1") +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "Practice 3M5",
       subtitle = sprintf("Grid size = %d", nrow(df)),
       x = "grid values of probability of water", 
       y = "posterior probability simulated")
```


Create the sampled data


```{r}
n_samples <- 1e4
set.seed(100)
the_samples <- df %>%
    slice_sample(n = n_samples, weight_by = posterior, replace = TRUE) %>%
    mutate(id = seq_len(n_samples), .before = p_grid)
```

and find the HPDI

```{r}
# HPDI with rethinking
rethinking::HPDI(the_samples$p_grid, prob = 0.9)
```

```{r}
# HPDI with ggdist
ggdist::mean_hdi(.data = the_samples$p_grid, .width = 0.9)
```


Constructing a posterior prediction by simulating $W$ with the sampled $p$.

```{r}
# using the sampled p we simulate the water
post_pred <- rbinom(n = nrow(the_samples), size = n_trials, prob = the_samples$p_grid)
the_samples <- data.frame(
  the_samples,
  predict = post_pred
)
```


the histogram of the predictions is

```{r}
colr <- paletteer::paletteer_d("Manu::Hoiho", direction = -1)
ggplot(the_samples, aes(x=predict)) +
  geom_histogram(aes(fill = ..count..)) +
  scale_x_continuous(breaks = 0:15) +
  scale_fill_gradientn(colors = colr) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  labs(title = "Practice 3M5",
       subtitle = sprintf("Frequencies of %d predictions", nrow(the_samples)))
```

## 3M6 {-#prac3M6}


We perform a sampling from a grid as in section 3.1 but, this time, with a grid
covering two variables: the *percentage of water (probability of success)* $p_water$ 
*as well as the nb of tosses (trials)* that is $n_tosses$.

The assumed prior for $p_water$ is flat.  You can change this at your discretion.

```{r}
p_water <- seq(from = 0, to = 1, by = 0.01)
n_tosses <- seq(from = 500, to = 5000, by = 500)
the_grid <- expand_grid(p_water, n_tosses)
# assume flat uniform prior, feel free to change it
the_grid$prior <- rep(1, times = nrow(the_grid))
stopifnot(nrow(the_grid) == length(p_water) * length(n_tosses))
```

and we compute the posterior probability of every $p_water$ in the grid, given 
$n_tosses$. Since we don't have $y$, that is the observed nb of successes,
we simulate them, assuming $p=0.7$ with `rbinom(x = 1, size = the_grid$n_tosses, prob = 0.7)`.

```{r}
# compute the x, nb of success using rbinom()
set.seed(3)
x_water <- rbinom(n = 1, size = the_grid$n_tosses, prob = 0.7)
likelihood <- dbinom(x = x_water, size = the_grid$n_tosses, prob = the_grid$p_water)
# compute the posterior of p
posterior <- likelihood * the_grid$prior
posterior <- posterior / sum(posterior)
the_grid$posterior <- posterior
# the posterior must add up to 1
stopifnot(sum(the_grid$posterior) == 1)
```

and we finally create a matrix of simulations where every column of the matrix
is a simulated sample using a given nb of tosses.

```{r}
# create matrix of samples, each column is an assumed nb of tosses
msamples <- sapply(X = n_tosses, FUN = function(x){
  a_grid <- the_grid[the_grid$n_tosses == x, ]
  samples <- sample(x = a_grid$p_water, size = 1000, prob = a_grid$posterior, replace = TRUE)
  return(samples)
})
# name the columns after the nb of tosses
colnames(msamples) <- n_tosses
```

and the final answer is given by finding the width of the 99% interval fro every
assumption of $n_tosses$ (column). The quantile interval `rethinking::PI(x, prob = 0.99)`
but is used, the high-density interval `rethinking::HPDI(x, prob = 0.99)` could also used.

```{r}
# find the width of the range for each sample (column)
(apply(X = msamples, MARGIN = 2, FUN = function(x) diff(rethinking::PI(x, prob = 0.99))))
```

> The answer is therefore that the nb of tosses should >= 2000, assuming
a flat uniform prior, using increments of 500 for the nb of tosses.


## 3H1 {-#prac3H1}

Loading the data from the rethinking package

```{r}
data(homeworkch3)
```


Get the prior estimate from the data and calculate the posterior

```{r}
n_boys <- sum(birth1) + sum(birth2)
n_births <- length(birth1) + length(birth2)
p_boys <- n_boys / n_births
p_grid <- seq(from = 0, to = 1, by = 0.01)
p_prior <- rep(1, times = length(p_grid))
df <- data.frame(
    p_grid = p_grid,
    prior = p_prior) %>%
    mutate(
     likelihood = dbinom(x = n_boys, size = n_births, prob = p_grid),
     posterior = (likelihood * prior) / sum(likelihood * prior)
    )
stopifnot(sum(df$posterior) == 1)
```

using the position of the maximum posterior prob. we find the p that maximize 
the posterior

```{r}
(df$p_grid[which.max(df$posterior)])
```

## 3H2 {-#prac3H2}

```{r}
samples <- sample(x = df$p_grid, size = 1e4, prob = df$posterior, replace = TRUE)
# 50%, 89%, 97% with HPDI
rethinking::HPDI(samples, prob = c(0.5, 0.89, 0.97))
```

and with `gdist`

```{r}
ggdist::mean_hdi(.data = samples, .width = c(0.50, 0.89, 0.97))
```


## 3H3 {-#prac3H3}

```{r}
# do a posterior prediction
pred <- data.frame(pred = 
                     rbinom(n = length(samples), size = n_births, prob = samples)
                   )
# avg nb of boys in the simulation
mean(pred$pred)
# using the HPDI to find the interval
rethinking::HPDI(pred$pred, prob = 0.89)
```

or getting the mean and interval all at once with `ggdist`

```{r}
ggdist::mean_hdi(.data = pred$pred, .width = 0.89)
```

```{r}
rng <- range(pred$pred)
qtl <- c(0.5, 0.8, 0.95, 1)
x_breaks <- ggdist::mean_qi(.data = pred$pred, .width = qtl) %>%
  select(y, ymin, ymax) %>%
  pivot_longer(cols = c("y", "ymin", "ymax")) %>%
  distinct(value) %>%
  arrange(value) %>%
  pull() %>%
  as.integer()

# the actual binomial distribution of the observations
set.seed(3)
dist_binom <- data.frame(x = rbinom(n = length(samples), size = n_births, prob = p_boys))
ggplot(pred, aes(x = pred)) +
  stat_halfeye(aes(fill=stat(cut_cdf_qi(cdf, .width = qtl))),
               point_interval = ggdist::mean_qi, .width = qtl) +
  geom_density(data = dist_binom, mapping = aes(x, y = ..scaled..), color = "violetred", linetype = "dashed", size = 1) +
  geom_vline(xintercept = c(mean(dist_binom$x), mean(pred$pred)), 
             color = c("violetred", "darkgreen"), linetype = c("dashed", "solid")) +
  scale_x_continuous(breaks = x_breaks,limits = rng) +
  scale_fill_paletteer_d(palette = "Manu::Takahe", direction = -1,
                         na.translate = FALSE) +
  theme_ggdist() +
  theme(legend.position = c(0.1,0.75),
        legend.title = element_text(color = "navy")) +
  labs(title = "Practice 3H3", x = "nb of boys", y = "probability", fill = "quantiles")
```

## 3H4 {-#prac3H4}

Finding the posterior dist with only the first-born

```{r}
n_boys <- sum(birth1)
n_births <- length(birth1)
p_boys <- n_boys / n_births
p_grid <- seq(from = 0, to = 1, by = 0.01)
p_prior <- rep(1, times = length(p_grid))
df <- data.frame(
    p_grid = p_grid,
    prior = p_prior) %>%
    mutate(
     likelihood = dbinom(x = n_boys, size = n_births, prob = p_grid),
     posterior = (likelihood * prior) / sum(likelihood * prior)
    )
stopifnot(sum(df$posterior) == 1)
```

and create a sample of the $p$ from the grid and its posterior.

```{r}
set.seed(3)
samples <- sample(x = df$p_grid, size = 1e4, prob = df$posterior, replace = TRUE)
```



```{r}
# do a posterior prediction
pred <- data.frame(pred = 
                     rbinom(n = length(samples), size = n_births, prob = samples)
                   )
# avg nb of boys in the simulation
mean(pred$pred)
# using the HPDI to find the interval of prior
rethinking::HPDI(pred$pred, prob = 0.89)
```

or getting the prior's mean and interval all at once with `ggdist`

```{r}
ggdist::mean_hdi(.data = pred$pred, .width = 0.89)
```

```{r}
rng <- range(pred$pred)
# the acutal binomial distribution of the observations
set.seed(3)
dist_binom <- data.frame(x = rbinom(n = length(samples), size = n_births, prob = p_boys))
ggplot(pred, aes(x = pred)) +
  stat_halfeye(point_interval = ggdist::mean_hdi, .width = 0.89, fill = "rosybrown1") +
  geom_density(data = dist_binom, mapping = aes(x, y = ..scaled..), color = "navy", linetype = "dashed") +
  geom_vline(xintercept = c(mean(dist_binom$x), mean(pred$pred)), 
             color = c("navy", "rosybrown4"), linetype = c("dashed", "solid")) +
  scale_x_continuous(limits = rng) +
  theme_minimal() +
  labs(title = "Practice 3H4", x = "nb of boys", y = "probability")
```


## 3H5 {-#prac3H5}

The idea here is that, if boys are independent of girls then
$P(boy \mid girl) = P(boy)$ when $boy$ and $girl$ are independent.


Finding the posterior dist with boys born after a girl.

```{r}
# births of girls followed by boy
girlboy <- (birth1 == 0) & (birth2 == 1)
n_girlboy <- sum(girlboy)
n_births <- length(girlboy)
p_girlboy <- n_girlboy / n_births
p_grid <- seq(from = 0, to = 1, by = 0.01)
p_prior <- rep(1, times = length(p_grid))
df <- data.frame(
    p_grid = p_grid,
    prior = p_prior) %>%
    mutate(
     likelihood = dbinom(x = n_girlboy, size = n_births, prob = p_grid),
     posterior = (likelihood * prior) / sum(likelihood * prior)
    )
stopifnot(sum(df$posterior) == 1)
```


Get a sample of $p$ from the grid using the posterior as probability of the given value.
Then make a prediciton of boys using these sampled $p$ values

```{r}
# sample the grid using the posterior probabilities
samples <- sample(x = df$p_grid, size = 10000, replace = TRUE, prob = df$posterior)
# simulate births
pred <- data.frame(pred = 
                     rbinom(n = length(samples), size = n_births, prob = samples)
                   )
# using the HPDI to find the interval
HPDI(pred$pred, prob = 0.89)
```

and visualize

```{r}
# draw a graph with the result
rng <- range(pred$pred)
# the acutal binomial distribution of the observations
set.seed(3)
dist_binom <- data.frame(x = rbinom(n = length(samples), size = n_births, prob = p_girlboy))
ggplot(pred, aes(x = pred)) +
  stat_halfeye(point_interval = ggdist::mean_hdi, .width = 0.89, fill = "peachpuff") +
  geom_density(data = dist_binom, mapping = aes(x, y = ..scaled..), color = "slateblue4", linetype = "dashed") +
  geom_vline(xintercept = c(mean(dist_binom$x), mean(pred$pred)), 
             color = c("slateblue4", "peachpuff4"), linetype = c("dashed", "solid")) +
  scale_x_continuous(limits = rng) +
  theme_minimal() +
  labs(title = "Practice 3H5", x = "nb of boys", y = "probability")
```

The average nb of first boys was about 50, see 3H4 above.
Now, using the second born boy given a girl was born, the average is 40.
It should remain about the same if boys were independent f girls and it
is obviously not.

Conclusion: Birth of boy is not independent of girls.
