```{r include=FALSE}
library(rethinking)
library(brms)
library(INLA)
library(dplyr, quietly = TRUE)
library(tidyr, quietly = TRUE)
library(tidybayes, quietly = TRUE)
library(modelr)
library(scales)
library(loo)
library(simstudy)
library(dagitty, quietly = TRUE)
library(ggdag, quietly = TRUE)
library(posterior)
library(ggdist, quietly = TRUE)
library(ggmcmc, quietly = TRUE)
library(bayesplot, quietly = TRUE)
library(patchwork, quietly = TRUE)
library(paletteer, quietly = TRUE)
library(gt)
```


# Monsters and Mixtures {#Mixed}

We set the current theme used for plotting

```{r}
theme_set(
  ggthemes::theme_hc(base_size = 12, base_family = "sans", style = "darkunica") +
  theme(title = element_text(color = "floralwhite"),
        axis.title.y = element_text(angle = 90),
        strip.background = element_rect(fill = "darkgrey"))
  )
```



## 12E1 {-#prac12E1}

See introduction to section 12.3, p. 380 for reference.

* *Unordered categorical variable*: Variable with discrete values such as 
categories of products or classes of clients.
* *Ordered categorical variable*: Variables with discrete values whose ranking
is meaningful but the distance between the different values is unequal. For
example the popularity of a product could be categorized as very low, low,
medium, high, very high.


## 12E2 {-#prac12E2}

See introduction to section 12.3, p. 380, for reference.


Use the *cumulative link* function. In differ from the logit link in that it 
represents the probability of the value *and any smaller values*.

That is, whereas the logit link is the probability of category 3 happening, the
cumulative link function is the probability of categories 1, 2 or 3 happening.


## 12E3 {-#prac12E3}

See introduction to section 12.2, p. 376, for reference.

Zero-inflated model are caused by 2 different factor: Either the process didn't 
generate and event because of a low rate or simply because the process failed
altogether.

Therefore the inferential error will be to understimate the process rate.


## 12E4 {-#prac12E4}


See introduction to section 12.1, p. 369-370, for reference.

Over-dispersion is when the values are more variable than the pure process. *It
implies that some omitted variable is producing additional variations*. For
example it happens often with sales data as some locations sell more than 
expected for undeclared reason.

Under-dispersion is when the values are less variable than the pure process.
For example it happens in produciton when the rate of production is more constant
when the volume is high, i.e. there are many jobs in the queue. This is described
by the *Conway-Maxwell-Poisson* distribution.


## 12M1 {-#prac12M1}

The log-cumulative-odds $\alpha_k$ is defined in equation 12.1, section 12.3.2, 
p. 383, as

$$
\begin{align*}
\alpha_k = log{\frac{Pr(y_i \le k)}{1 - Pr(y_i \le k)}}
\end{align*}
$$

see R. code 12.15 on p. 383 to see an example on how to do this. The last
value is `Inf` since probability of the last value is always 1.

```{r}
x <- c(12, 36, 7, 41)  # the counts
probs <- cumsum(x) / sum(x)  # the cumulative probabilities
odds <- probs / (1 - probs)  # the cumulative odds
cumlogit <- log(odds)  # the log of cumulative odds
cumlogit
```

## 12M2 {-#prac12M2}


Figure 12.5 is in 12.3.1, p. 384.


```{r out.width="80%"}
p <- list()
p$df <- data.frame(response = 1:4,
                   cnt = x,
                   probs = probs,
                   diff = x / sum(x),
                   previous = c(0, probs[-4]))
ggplot(p$df, aes(x = response, y = probs)) +
  geom_linerange(aes(x = response, ymin = 0, ymax = probs), 
                 color = "indianred1", size = 3) +
  geom_linerange(aes(x = response + 0.05, ymin = previous, ymax = probs), 
                 color = "darkorange", size = 3) +
  geom_line(color = "lightcoral", size = 1) +
  geom_point(color = "lightblue", size = 4, shape = 19) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(title = "Cumulative probabilities and ordered likelihood",
       subtitle = "Practice 12M2", y = "cumulative proportion")
```

## 12M3 {-#prac12M3}


See section 12.2.1 for Zero-inflated Poisson.

The Zero-inflated binomial would be using the same pattern with $p_0$ the 
probability that the process did not start and thus we have a zero.

We also have $p_y$ which is the probability of success when the process operates
normally and we have $n$ trials.

So the probability of obtaining a zero is $p_0$ + the probability of normally
a zero with the process is $(1-p_0)(1-p)^n$, that is

$$
Pr(y = 0 \mid p_0, p_y, n) = p_0 + (1-p_0) \cdot (1-p_y)^n
$$
And the probability of obtaining $y_i = k, k \gt 0$ depends on the probability of the
process not failing $1-p_0$ and the usual binomial probability of the process
$\binom{n}{k}(1-p_y)^{n-k}p_y^k$, therefore

$$
Pr(y = k \mid p_0, p_y, n) = (1-p_0) \cdot \binom{n}{k}(1-p_y)^{n-k}p_y^k
$$

which can be formulated as a model as follows


$$
\begin{align*}
y_i &\sim \mathcal{ZIBinomial(p_{0,i}, p_{y,i})} \\
logit(p_{0,i}) &= \alpha_{p0} + \beta_{p0}x_i \\
logit(p_{y,i}) &= \alpha_{py} + \beta_{py}x_i
\end{align*}
$$


## 12H1 {-#prac12H1}

We don't transform $femiminity$ as its scale is from 1 to 10 and no bias
can be caused by large predictors's values.


```{r}
data("Hurricanes")
dataHurricanes <- Hurricanes %>%
  mutate(pressure = rescale(min_pressure, to = c(1, 10)),
         damage = log(damage_norm))
skimr::skim(dataHurricanes)
rm(Hurricanes)
```

The transformations are as follows

* $pressure = rescale(min\_pressure, c(1, 10))$: The $min\_pressure$ rescaled 
to `c(1, 10)` since
  * The pressure is always strictly positive
  * Avoid the distortion of coefficient with the high observed range
  * `c(1, 10)` is a scale similar to $femininity$ and facilitate the comparison
* $damage = log(damage\_norm)$: The log of $damage\_norm$ since
  * The distribution is heavily right-skewed and the log will centralize it
  * The lower value being 1 the log will give a non-negative scale which simplifies
  * It puts the data on a scale similar to $femininity$ which facilitates comparison


which we can illustrate as follows

```{r}
ggplot(dataHurricanes, aes(x = femininity, y = deaths)) +
  geom_point(color = "ivory") +
  ggrepel::geom_text_repel(data = dataHurricanes %>% filter(deaths >= 50),
                           aes(x = femininity, y = deaths, label = name),
                           color = "gold") +
  scale_y_continuous(breaks = scales::breaks_width(width = 50)) +
  labs(title = "Nb of deaths vs femininity",
       subtitle = sprintf("%d storms", nrow(dataHurricanes)))
```



### a) Femininity {-}

#### Model {-}

McElreath requires the **simplest Poisson model** of $deaths$ using $feminity$ 
as a predictor.


$$
\begin{align*}
deaths_i &\sim \mathcal{Poisson(\lambda_i)} \\
\log{\lambda_i} &= \alpha + \beta_i femininity_i \\
\alpha &\sim \mathcal{N(mean=4, sd=1)} \\
\beta_i &\sim \mathcal{N(mean=1, sd=0.25)}
\end{align*}
$$


#### Priors {-}


We note that since $\log{\lambda_i} &= \alpha$ then $\lambda_i$ is lognormal
distributed and we could have use the $rlnorm()$ function for sampling.
We choose to use `simstudy` for consistency with part a of this practice

"Prior predictive distribution of mean deaths (lambda)"

```{r}
sim <- list(nsamples = 100)
sim <- within(sim, {
  # define the model
  defs <- defData(varname = "a", dist = "normal", formula = "..m",
                  variance = "..v")
  defs <- defData(defs, varname = "lambda", dist = "nonrandom", 
                  formula = "a", link = "log")
  
  # generate data using the grid f specs
  grid <- expand.grid(mean = c(2, 3, 4), 
                      sd = c(0.5, 1, 2)) %>%
    mutate(model = sprintf("m=%.1f, s=%.1f", mean, sd))
  
  lst <- lapply(seq_len(nrow(grid)), FUN = function(i) {
    m <- grid$mean[i]
    s <- grid$sd[i]
    v <- s^2
    set.seed(as.integer(as.Date("2021-12-27")))
    genData(n = nsamples, dtDefs = defs) %>%
      as.data.frame() %>%
      mutate(mean = paste("mu == ", m), 
             sd = paste("sigma == ", s), 
             model = grid$model[i])
    })
  names(lst) <- grid$model
  
  data <- do.call(rbind, lst)
})
# glimpse(sim$data)

p <- list()
p$title <- expression(paste('Prior predictive of ', lambda *': '* 
                              log(lambda) == alpha *', '* alpha %~% 
                              Normal(mu *', '* sigma)))
p$plot <- ggplot(sim$data, aes(x = lambda, fill = as.factor(mean), color = as.factor(mean))) +
  geom_density(aes(y = ..scaled..)) +
  scale_fill_paletteer_d("khroma::vibrant") +
  scale_color_paletteer_d("khroma::vibrant") +
  coord_cartesian(xlim = c(0, 100)) +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(title = p$title, x = NULL, y = NULL) +
  facet_grid(sd ~ mean, labeller = label_parsed)
p$plot
```


We choose to use $\alpha \sim \mathcal{N(mean=4, sd=1)}$ as prior.

See section 11.2., p. 351, on details about the priors for Poisson.


```{r}
sim <- list(nsamples = 25, npreds = 25)
sim <- within(sim, {
  defs = defData(varname = "a", dist = "normal", formula = 4, variance = 1^2)
  defs <- defData(defs, varname = "b", dist = "normal",
                  formula = "..m", variance = "..v")
  grid <- expand.grid(mean = c(0.5, 1, 1.5), sd = c(0.10, 0.25, 0.5)) %>%
    mutate(params = paste("mu ==", mean, "*','* ~ sigma ==", sd))
  lst <- lapply(X = seq_len(nrow(grid)), FUN = function(i) {
    prm <- grid$params[i]
    m <- grid$mean[i]
    s <- grid$sd[i]
    v <- s^2
    set.seed(as.integer(as.Date("2021-12-24")))
    genData(n = nsamples, dtDefs = defs) %>%
      tidyr::expand(nesting(id, a, b, m, s),
                    x = seq_range(x = log(dataHurricanes$femininity), n = npreds)) %>%
      mutate(lambda = exp(a + b * x),
             params = prm)
    })
  names(lst) <- grid$params

  data <- do.call(rbind, lst)
  })
# glimpse(sim$data)


p <- list()
p$y_lim <- c(0, 500)
p$h_line <- 300
p$pal <- paletteer::paletteer_d("fishualize::Hypsypops_rubicundus")
p$pal <- scales::col_factor(palette = unclass(p$pal),
                            domain = sort(unique(sim$data$params)),
                            na.color = "grey")
p$pal <- p$pal(unique(sim$data$params))
names(p$pal) <- unique(sim$data$params)
# p$pal
p$log <- ggplot(sim$data, aes(x = x, y = lambda, group = id, color = params)) +
  geom_line() +
  scale_fill_manual(values = p$pal) +
  scale_color_manual(values = p$pal) +
  coord_cartesian(ylim = p$y_lim) +
  theme(legend.position = "none") +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(x = "log of feminity", y = expression(paste("mean (", lambda, ") deaths"))) +
  facet_wrap(. ~ params, labeller = label_parsed)
# p$log

p$nat <- ggplot(sim$data, aes(x = exp(x), y = lambda, group = id, color = params)) +
  geom_line() +
  geom_hline(yintercept = p$h_line, size = 1, linetype = "dashed", color = "green") +
  # scale_fill_manual(values = p$pal) +
  scale_color_manual(values = p$pal) +
  coord_cartesian(ylim = p$y_lim) +
  theme(legend.position = "none") +
  # theme(legend.position = "none",
  #       axis.text.y = element_blank(),
  #       axis.ticks.y = element_blank()) +
  labs(x = "feminity", y = expression(paste("mean (", lambda, ") deaths"))) +
  facet_wrap(. ~ params, labeller = label_parsed)
# p$nat
wrap_plots(p$nat, p$log) +
  plot_annotation(title = expression(paste("Prior distribution of ", lambda)))
```

We select $\beta_i \sim \mathcal{N(mean=1.5, sd=0.25)}$ as the prior.


#### Fits {-}


##### `quap` {-}


```{r}
a_file <- here::here("fits", "m12H01a.rds")
m12H01a <- readRDS(file = a_file)
# m12H01a <- quap(
#   data = dataHurricanes,
#   flist = alist(
#     deaths ~ dpois(lambda),
#     log(lambda) <- a + b * femininity,
#     a ~ dnorm(4, 1),
#     b ~ dnorm(1.5, 0.25))
#   )
# saveRDS(m12H01a, file = a_file)
```

the summary is

```{r}
precis(m12H01a)
```




##### `brm` {-}


```{r}
a_file <- here::here("fits", "b12H01a.rds")
b12H01a <- readRDS(file = a_file)
# b12H01a <- brm(data = dataHurricanes,
#                    family = poisson,
#                    formula = deaths ~ 1 + femininity,
#                    prior = c(prior(normal(4, 1), class = Intercept),
#                              prior(normal(1, 0.25), class = b)),
#                    cores = detectCores())
# b12H01a <- brms::add_criterion(b12H01a, criterion = c("waic", "loo"))
saveRDS(b12H01a, file = a_file)
```

the summary is

```{r}
summary(b12H01a)
```



##### `inla` {-}



```{r}
a_file <- here::here("fits", "i12H01a.rds")
i12H01a <- readRDS(file = a_file)
# i12H01a <- inla(data = dataHurricanes,
#                 formula =  deaths ~ femininity,
#                 family = "poisson",
#                 control.family = list(link = "log"),
#                 control.fixed = list(mean.intercept = 4, prec.intercept = 1 / 1^2,
#                                      mean = 1, prec = 1 / 0.25^2),
#                 control.predictor = list(link = 1, compute = TRUE),
#                 control.compute = list(dic = TRUE, cpo = TRUE, waic = TRUE))
# saveRDS(i12H01a, file = a_file)
```

and the summary is

```{r}
summary(i12H01a)
```



### b) Intercept only {-}


#### Model {-}


$$
\begin{align*}
deaths_i &\sim \mathcal{Poisson(\lambda_i)} \\
\log{\lambda_i} &= \alpha \\
\alpha &\sim \mathcal{N(mean=4, sd=1)}
\end{align*}
$$


#### Fits {-}


##### `quap` {-}


```{r}
a_file <- here::here("fits", "m12H01b.rds")
m12H01b <- readRDS(file = a_file)
# m12H01b <- quap(
#   data = dataHurricanes,
#   flist = alist(
#     deaths ~ dpois(lambda),
#     log(lambda) <- a,
#     a ~ dnorm(4, 1))
#   )
# saveRDS(m12H01b, file = a_file)
```

the summary

```{r}
summary(m12H01b)
```



##### `brm` {-}


```{r}
a_file <- here::here("fits", "b12H01b.rds")
b12H01b <- readRDS(file = a_file)
# b12H01b <- brm(data = dataHurricanes,
#                    family = poisson,
#                    formula = deaths ~ 1,
#                    prior = prior(normal(4, 1), class = Intercept),
#                    cores = detectCores())
# b12H01b <- brms::add_criterion(b12H01b, criterion = c("waic", "loo"))
# saveRDS(b12H01b, file = a_file)
```

the summary

```{r}
summary(b12H01b)
```



##### `inla` {-}



```{r}
a_file <- here::here("fits", "i12H01b.rds")
i12H01b <- readRDS(file = a_file)
# i12H01b <- inla(data = dataHurricanes,
#                 formula =  deaths ~ 1,
#                 family = "poisson",
#                 control.family = list(link = "log"),
#                 control.fixed = list(mean.intercept = 4, prec.intercept = (1 / 1^2)),
#                 control.predictor = list(link = 1, compute = TRUE),
#                 control.compute = list(dic = TRUE, cpo = TRUE, waic = TRUE))
# saveRDS(i12H01b, file = a_file)
```

and the summary is

```{r}
summary(i12H01b)
```


### Inferences {-}


**How strong is femininity relation to death?**


```{r}
print(loo_compare(b12H01a, b12H01b), simplify = FALSE)
```

The model `b12H01b` which is uses femininity as a predictor seems to be a better 
fit. However the elpd_diff of -23.5 has a sd of 76.2 whihc is more than
3 times the coefficient itself and therefore uncertain.


**Which storms does it fit poorly?**

Th fit seems to be particularly poor for Camille, Diane, Ike and Sandy as shown 
when we plot the WAIC penalty vs the pareto K measure from loo.


```{r}
p <- list()
p$df <- data.frame(paretoK = b12H01a$criteria$loo$diagnostics$pareto_k,
                   penaltyWAIC = b12H01a$criteria$waic$pointwise[, "p_waic"],
                   name = pull(dataHurricanes, name),
                   deaths = pull(dataHurricanes, deaths)) %>%
  mutate(out = paretoK >= 0.5)
p$out <- ggplot(p$df, aes(x = paretoK, y = penaltyWAIC, color = out)) +
  geom_point() +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "violetred", size = 1) +
  ggrepel::geom_text_repel(aes(label = name), size = 4) +
  scale_x_continuous(breaks = scales::breaks_width(width = 0.25)) +
  scale_color_manual(values = c("TRUE" = "violetred", "FALSE" = "royalblue")) +
  theme(legend.position = "none") +
  labs(title = "Outlier diagnostic", subtitle = "Fit b12H01a")
p$out
```



## 12H2 {-#prac12H2}


### Model {-}


$$
\begin{align*}
deaths_i &\sim \mathcal{Gamma-Poisson(\lambda_i, \phi)} \\
\log{\lambda_i} &= \beta_i femininity_i \\
\beta_i &\sim \mathcal{N(mean=3, sd=0.50)} \\
\phi &\sim \mathcal{Exp(rate = 1)}
\end{align*}
$$


### Priors {-}

We use the same prior for $\beta_i$ as in the previous practice 12H1.  Since $\phi$
represents the variance of the Gamma-Poisson (p. 374 of section 12.1.2) then
we use $\phi \sim \mathcal{Exp(rate = 1)}$.


### Fits {-}


#### `quap` {-}

See the documentation of `rethinking::dgampois()` which uses this practice as
an example, I use different priors (see 12H1 for details)

See [more rethinking](http://xcelab.net/R/rethinking_package.pdf) for additional
details on how to use rethinking.


```{r}
a_file <- here::here("fits", "m12H02.rds")
m12H02 <- readRDS(file = a_file)
# m12H02 <- quap(
#   data = dataHurricanes,
#   flist = alist(
#     deaths ~ dgampois(lambda, phi),
#     log(lambda) <- a + b * femininity,
#     a ~ dnorm(4, 1),
#     b ~ dnorm(1, 0.25),
#     phi ~ dexp(1))
#   )
# saveRDS(m12H02, file = a_file)
```

the summary

```{r}
precis(m12H02)
```


#### `brm` {-}



```{r}
a_file <- here::here("fits", "b12H02.rds")
b12H02 <- readRDS(file = a_file)
# b12H02 <- brm(data = dataHurricanes,
#       family = negbinomial,
#       deaths ~ 1 + femininity,
#       prior = c(prior(normal(4, 1), class = Intercept),
#                 prior(normal(1, 0.25), class = b),
#                 prior(gamma(0.01, 0.01), class = shape)),
#       cores = detectCores(),
#       seed = 12)
# # NOTE: be careful when using waic with gamma-poisson
# b12H02 <- brms::add_criterion(b12H02, criterion = c("waic", "loo"))
# saveRDS(object = b12H02, file = a_file)
```

the summary

```{r}
summary(b12H02)
```

### Inference {-}

The posterior distribution of the coefficients using the model `b12H01a`
(Poisson) and `b12H02` (Poisson-Gamma) is

```{r}
samples <- list(probs = c(0.055, 0.945))
samples <- within(samples, {
  x <- seq_range(dataHurricanes$femininity, n = 50, pretty = TRUE)
  pois <- as_draws_df(b12H01a) %>%
    summarise_draws(mean, median, mode = ggdist::Mode, sd, mad, 
                    ~quantile2(., probs = probs)) %>%
    mutate(fit = "b12H01a Poisson") %>%
    filter(variable != "lp__")
  pgam <- as_draws_df(b12H02) %>%
    summarise_draws(mean, median, mode = ggdist::Mode, sd, mad, 
                    ~quantile2(., probs = probs)) %>%
    mutate(fit = "b12H02 Poisson-Gamma") %>%
    filter(!(variable %in% c("shape", "lp__")))
  data <- rbind(pois, pgam)
})
# glimpse(samples$pois)
p <- list()
p$coefs <- ggplot(samples$data, aes(x = mean, xmin = q5.5, xmax = q94.5, 
                                    y = variable, color = fit)) +
  geom_pointinterval(fatten_point = 3, size = 2,
                     position = position_dodge(width = 0.5)) +
  geom_vline(xintercept = 0, color = "white", linetype = "dashed") +
  scale_x_continuous(breaks = scales::breaks_extended(n = 7)) +
  theme(legend.title = element_blank()) +
  labs(title = "Posterior 89% range of coefficients",
        x = NULL, y = NULL)
p$coefs
```


We can see that the coefficients, in particular the $\beta_i$ is more variable
with the gamma-poisson fit than the poisson fit.

This is caused by the fact that the Poisson fit has a standard deviation
equal to its mean whereas the Gamma-Poisson fit allows the standard deviation
$\sigma$ to using the gamma distribution. See the point made by McElreath
in the last sentence of the first paragraph of section 12.1.2 on p. 373.


We can also show the same result with the predictive distribution

```{r}
predict <- list(probs = c(0.055, 0.945))
predict <- within(predict, {
  x <- data.frame(femininity = seq_range(dataHurricanes$femininity, n = 50, pretty = TRUE))
  pois <- predicted_draws(b12H01a, newdata = x) %>%
    ungroup() %>%
    select(-.row, -.chain, -.iteration, -.draw) %>%
    summarise_draws(mean, median, sd, mad,
                    ~quantile2(., probs = probs)) %>%
    filter(variable == ".prediction") %>%
    mutate(fit = "b12H01a Poisson")
    # filter(variable != "lp__")
  pgam <- predicted_draws(b12H02, newdata = x) %>%
    ungroup() %>%
    select(-.row, -.chain, -.iteration, -.draw) %>%
    summarise_draws(mean, median, sd, mad,
                    ~quantile2(., probs = probs)) %>%
    filter(variable == ".prediction") %>%
    mutate(fit = "b12H02 Gamma-Poisson")
  data <- rbind(pois, pgam)
})
# glimpse(predict$pois)
p <- list()
p$coefs <- ggplot(predict$data, aes(x = mean, xmin = q5.5, xmax = q94.5, 
                                    y = variable, color = fit)) +
  geom_pointinterval(fatten_point = 5, size = 3,
                     position = position_dodge(width = 0.5)) +
  geom_vline(xintercept = 0, color = "white", linetype = "dashed") +
  scale_x_continuous(breaks = scales::breaks_extended(n = 5)) +
  theme(legend.title = element_blank()) +
  labs(title = "Posterior 89% range of predicted deaths",
        x = NULL, y = NULL)
p$coefs
```



## 12H3 {-#prac12H3}

As mentioned in practice 2H1 *the explanation given is that people
unconsciously rate female hurricanes as less dangerous and so are less likely
to evacuate* which increases the probability of deaths.

### DAG {-}

The structural causal model is build on the following assumptions

* $pressure$ cannot cause $deaths$ by itself but through $damage$ such as falling debris
or $femininity$ because people are less likely to evacuate with a more feminine name. 
Hence $pressure$ is a cause of both $damage$ and $femininity$.
* $damage$ is a cause of deaths (e.g. falling/flying debris)
* The $femininity$ of the name is a cause of $deaths$ because people are less
likely to evacuate (see 12H1)
* $damage$ is known **after** $femininity$ and therefore $damage$ cannot
possibly influence $femininity$
* $femininity$ is known **before** $damage$ and, thus, could presumably cause it


```{r}
scm <- list()
scm <- within(scm, {
  dag <- dagify(
  deaths ~ damage + femininity,
  damage ~ pressure + femininity,
  femininity ~ pressure,
  outcome = "deaths")
  
  p <- dag %>% 
    tidy_dagitty(layout = "sugiyama") %>%
    ggdag::ggdag(node = FALSE) +
    geom_dag_text(size = 5, color = "tomato4") +
    geom_dag_edges(edge_color = "tomato4") +
    ggdag::theme_dag(panel.background = element_rect(fill = "cornsilk", color = "cornsilk")) +
    theme(title = element_text(color = "midnightblue")) +
    labs(title = "Structural causal model for Hurricanes",
         subtitle = "Practice 12H3")
})
scm$p
```

the conditional independences are

```{r}
impliedConditionalIndependencies(scm$dag)
```

### Priors {-}

Since we used the transformed variable $pressure$ and $damage$ which are on
a scale `c(1, 10)` similar to $femininity$ then we use the same priors
for all variables as in practice 12H1.

### Fits {-}

```{r}
# femininity
a_file <- here::here("fits", "b12H03_f.rds")
b12H03_f <- readRDS(file = a_file)
# b12H03_f <- brm(data = dataHurricanes,
#       family = negbinomial,
#       deaths ~ 1 + femininity,
#       prior = c(prior(normal(4, 1), class = Intercept),
#                 prior(normal(1, 0.25), class = b),
#                 prior(gamma(0.01, 0.01), class = shape)),
#       cores = detectCores(),
#       seed = 12)
# # NOTE: be careful when using waic with gamma-poisson
# b12H03_f <- brms::add_criterion(b12H03_f, criterion = c("waic", "loo"))
# saveRDS(object = b12H03_f, file = a_file)

# damage
a_file <- here::here("fits", "b12H03_d.rds")
b12H03_d <- readRDS(file = a_file)
# b12H03_d <- brm(data = dataHurricanes,
#       family = negbinomial,
#       deaths ~ 1 + damage,
#       prior = c(prior(normal(4, 1), class = Intercept),
#                 prior(normal(1, 0.25), class = b),
#                 prior(gamma(0.01, 0.01), class = shape)),
#       cores = detectCores(),
#       seed = 12)
# # NOTE: be careful when using waic with gamma-poisson
# b12H03_d <- brms::add_criterion(b12H03_d, criterion = c("waic", "loo"))
# saveRDS(object = b12H03_d, file = a_file)

# pressure
a_file <- here::here("fits", "b12H03_p.rds")
b12H03_p <- readRDS(file = a_file)
# b12H03_p <- brm(data = dataHurricanes,
#       family = negbinomial,
#       deaths ~ 1 + pressure,
#       prior = c(prior(normal(4, 1), class = Intercept),
#                 prior(normal(1, 0.25), class = b),
#                 prior(gamma(0.01, 0.01), class = shape)),
#       cores = detectCores(),
#       seed = 12)
# # NOTE: be careful when using waic with gamma-poisson
# b12H03_p <- brms::add_criterion(b12H03_p, criterion = c("waic", "loo"))
# saveRDS(object = b12H03_p, file = a_file)

# femininity and damage
a_file <- here::here("fits", "b12H03_fd.rds")
b12H03_fd <- readRDS(file = a_file)
# b12H03_fd <- brm(data = dataHurricanes,
#       family = negbinomial,
#       deaths ~ 1 + femininity + damage,
#       prior = c(prior(normal(4, 1), class = Intercept),
#                 prior(normal(1, 0.25), class = b),
#                 prior(gamma(0.01, 0.01), class = shape)),
#       cores = detectCores(),
#       seed = 12)
# # NOTE: be careful when using waic with gamma-poisson
# b12H03_fd <- brms::add_criterion(b12H03_fd, criterion = c("waic", "loo"))
# saveRDS(object = b12H03_fd, file = a_file)

# femininity and pressure
a_file <- here::here("fits", "b12H03_fp.rds")
b12H03_fp <- readRDS(file = a_file)
# b12H03_fp <- brm(data = dataHurricanes,
#       family = negbinomial,
#       deaths ~ 1 + femininity + pressure,
#       prior = c(prior(normal(4, 1), class = Intercept),
#                 prior(normal(1, 0.25), class = b),
#                 prior(gamma(0.01, 0.01), class = shape)),
#       cores = detectCores(),
#       seed = 12)
# # NOTE: be careful when using waic with gamma-poisson
# b12H03_fp <- brms::add_criterion(b12H03_fp, criterion = c("waic", "loo"))
# saveRDS(object = b12H03_fp, file = a_file)

# damage and pressure
a_file <- here::here("fits", "b12H03_dp.rds")
b12H03_dp <- readRDS(file = a_file)
# b12H03_dp <- brm(data = dataHurricanes,
#       family = negbinomial,
#       deaths ~ 1 + damage + pressure,
#       prior = c(prior(normal(4, 1), class = Intercept),
#                 prior(normal(1, 0.25), class = b),
#                 prior(gamma(0.01, 0.01), class = shape)),
#       cores = detectCores(),
#       seed = 12)
# # NOTE: be careful when using waic with gamma-poisson
# b12H03_dp <- brms::add_criterion(b12H03_dp, criterion = c("waic", "loo"))
# saveRDS(object = b12H03_dp, file = a_file)

# femininity, damage and pressure
a_file <- here::here("fits", "b12H03_fdp.rds")
b12H03_fdp <- readRDS(file = a_file)
# b12H03_fdp <- brm(data = dataHurricanes,
#       family = negbinomial,
#       deaths ~ 1 + femininity + damage + pressure,
#       prior = c(prior(normal(4, 1), class = Intercept),
#                 prior(normal(1, 0.25), class = b),
#                 prior(gamma(0.01, 0.01), class = shape)),
#       cores = detectCores(),
#       seed = 12)
# # NOTE: be careful when using waic with gamma-poisson
# b12H03_fdp <- brms::add_criterion(b12H03_fdp, criterion = c("waic", "loo"))
# saveRDS(object = b12H03_fdp, file = a_file)
```


### Inference {-}

get the posterior intervals

```{r}
samples <- list(probs = c(0.055, 0.945))

samples$fits <- list(
  "fmnn" = b12H03_f,
  "damg" = b12H03_d,
  "prss" = b12H03_p,
  "fmnn-damg" = b12H03_fd,
  "fmnn-prss" = b12H03_fp,
  "damg-prss" = b12H03_dp,
  "fmnn-damg-prss" = b12H03_fdp
)

samples$data <- purrr::map_dfr(.x = samples$fits, .f = function(x) {
  as_draws(x) %>%
    summarise_draws(mean, median, ~quantile2(., probs = samples$probs)) %>%
    filter(!(variable %in% c("lp__", "shape"))) %>%
    mutate(across(.cols = where(is.numeric), .fns = round, 4))
  },
  .id = "fit"
  )
# glimpse(samples$data)
```


when comparing the fits we have

```{r}
print(loo_compare(b12H03_f, b12H03_d, b12H03_p, b12H03_fd, b12H03_fp, 
            b12H03_dp, b12H03_fdp, criterion = "waic"),
      simplify = FALSE)
```

```{r}
round(model_weights(b12H03_f, b12H03_d, b12H03_p, b12H03_fd, b12H03_fp, 
            b12H03_dp, b12H03_fdp), 3)
```




```{r}
p <- list()
p$coefs <- ggplot(samples$data, aes(x = mean, xmin = q5.5, xmax = q94.5, 
                                    y = variable, color = fit)) +
  geom_pointinterval(fatten_point = 3, size = 2,
                     position = position_dodge(width = 0.5)) +
  geom_vline(xintercept = 0, color = "white", linetype = "dashed") +
  scale_x_continuous(breaks = scales::breaks_extended(n = 7)) +
  scale_color_paletteer_d("awtools::ppalette") +
  theme(legend.title = element_blank()) +
  labs(title = "Posterior distribution of coefficients with 89% interval",
        x = NULL, y = NULL)
p$coefs
```

We observe that

* one group which always include the damage and has a negative 
intercept on deaths but with much variability
* one group which seems to be mostly influenced, strongly, by pressure with
a positive intercept
* when pressure and damage are used together, the variability is very large
* femininity seems to have little to no impact on the patterns observed

We note that damage and pressure create 2 clusters of coefficients and that
they seem to have col linearity between them which causes large variability
when used together.

In conclusion it seems that the structural causal model above is justified
and that, in terms of actual effect, damage and pressure are much more important
than femininity which has little influence, if any, on the death rate.


The best fit seems to be with damage when using waic. The worst one is with
femininity!

when comparing the fits we have

```{r}
print(loo_compare(b12H03_f, b12H03_d, b12H03_p, b12H03_fd, b12H03_fp, 
            b12H03_dp, b12H03_fdp, criterion = "waic"),
      simplify = FALSE)
```

The weight attributed to `b12H03_d` is 87.5%.  Again, damage sems to be the
most important predictor of deaths.

```{r}
# takes a few seconds to run
# round(model_weights(b12H03_f, b12H03_d, b12H03_p, b12H03_fd, b12H03_fp, 
#             b12H03_dp, b12H03_fdp), 3)
```





## 12H4 {-#prac12H4}



