```{r include=FALSE}
library(rethinking)
library(brms)
library(INLA)
library(qs)
library(dplyr, quietly = TRUE)
library(tidyr, quietly = TRUE)
library(tidybayes, quietly = TRUE)
library(modelr)
library(scales)
library(loo)
library(simstudy)
library(dagitty, quietly = TRUE)
library(ggdag, quietly = TRUE)
library(posterior)
library(ggdist, quietly = TRUE)
library(ggmcmc, quietly = TRUE)
library(bayesplot, quietly = TRUE)
library(patchwork, quietly = TRUE)
library(paletteer, quietly = TRUE)
library(gt)
```



```{r}
# these options help stan run faster
# source: http://mjskay.github.io/tidybayes/articles/tidy-brms.html
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```



# Multilevel Models {#MLM}

We set the current theme used for plotting

```{r}
theme_set(ggthemes::theme_solarized_2())
```



## 13E1 {-#prac13E1}

$\alpha_{tank} \ sim \mathcal{N}(0, 1)$ should produce more shrinkage because
$\sigma = 1$ will force a smaller estimate than with $\sigma = 2$.


## 13E2 {-#prac13E2}

See section 13.1, p. 403 for example.

$$
\begin{align*}
y_i &\sim \mathcal{Binomial}(1, p_i) \\
logit(p_i) &= \alpha_{group[i]} + \beta x_i \\
\alpha_{j} &\sim \mathcal{N}(\bar{\alpha}, \sigma_{\alpha}) \\
\beta &\sim \mathcal{N}(0, 0.5) \\
\bar{\alpha} &\sim \mathcal{N}(0, 1.5) \\
\sigma_{\alpha} &\sim \mathcal{Exponential}(1)
\end{align*}
$$


## 13E3 {-#prac13E3}

See section 13.1, p. 403 for example.

$$
\begin{align*}
y_i &\sim \mathcal{N}(\mu, \sigma) \\
\mu_i &= \alpha_{group[i]} + \beta x_i \\
\alpha_{j} &\sim \mathcal{N}(\bar{\alpha}, \sigma_{\alpha}) \\
\beta &\sim \mathcal{N}(0, 1) \\
\bar{\alpha} &\sim \mathcal{N}(0, 5) \\
\sigma_{\alpha} &\sim \mathcal{Exp}(1) \\
\sigma &\sim \mathcal{Exp}(1)
\end{align*}
$$


## 13E4 {-#prac13E4}


See section 11.2 for Poisson regression.

$$
\begin{align*}
y_i &\sim \mathcal{Poisson}(\lambda_i) \\
\log{\lambda_i} &= \alpha_{group[i]} \\
\alpha_j &\sim \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \\
\bar{\alpha} &\sim \mathcal{N}(0, 5) \\
\sigma_{\alpha} &\sim \mathcal{Exp}(1) \\
\end{align*}
$$


## 13E5 {-#prac13E5}


See rethinking box at the beginning of section 13.3 on cross-classification.

$$
\begin{align*}
y_i &\sim \mathcal{Poisson}(\lambda_i) \\
\log{\lambda_i} &= \alpha_{group[i]} + \beta_{type[i]} \\
\alpha_j &\sim \mathcal{N}(\bar{\alpha}, \sigma_\alpha) \\
\bar{\alpha} &\sim \mathcal{N}(0, 5) \\
\sigma_{\alpha} &\sim \mathcal{Exp}(1) \\
\beta_j &\sim \mathcal{N}(\bar{\beta}, \sigma_\beta) \\
\bar{\beta} &\sim \mathcal{N}(0, 5) \\
\sigma_{\beta} &\sim \mathcal{Exp}(1) \\
\end{align*}
$$



## 13M1 {-#prac13M1}

The problem mentions $size$ and $pred$ as *effects* to add to the intercept.
Therefore they must be considered as predictors with numerical values and are
converted to dummy variables


```{r}
data(reedfrogs)
dataFrogs <- reedfrogs %>%
  mutate(tank = factor(seq_len(n())),
         pred = if_else(pred == "pred", 1, 0),
         size = if_else(size == "big", 1, 0))
rm(reedfrogs)
skimr::skim(dataFrogs)
```

and the model would be


$$
\begin{align*}
surv_i &\sim \mathcal{Binomial}(n_i, p_i) \\
logit(p_i) &= \alpha_{tank[i]} + \beta_{pred} \cdot pred_i + \beta_{size} \cdot size_i \\
\alpha_j &\sim \mathcal{N}(\bar{\alpha}, \sigma) \\
\bar{\alpha} &\sim \mathcal{N}(0, 1.5) \\
\beta_{pred}, \beta_{size} &\sim \mathcal{N}(0, 0.5)\\
\sigma &\sim \mathcal{Exponential}(1)
\end{align*}
$$

The problem mentions $size$ and $pred$ as *effects* to add to the intercept.
Therefore they must be considered as predictors with numerical values.

```{r}
a_file <- here::here("fits", "b13M01a.qs")
b13M01a <- qread(a_file)
# b13M01a <- brm(
#   data = dataFrogs,
#   family = binomial,
#   surv | trials(density) ~ 1 + (1 | tank),
#   prior = c(prior(normal(0, 1.5), class = Intercept),
#             prior(exponential(1), class = sd)),
#   cores = detectCores(), seed = 13)
# b13M01a <- add_criterion(b13M01a, c("loo", "waic"))
# qsave(b13M01a, a_file)

# cretae the list holding the fits
models <- list()
models <- within(models, {
  formulas <- list(
    "b13M01b" = bf(surv | trials(density) ~ 1 + (1 | tank) + pred),
    "b13M01c" = bf(surv | trials(density) ~ 1 + (1 | tank) + size),
    "b13M01d" = bf(surv | trials(density) ~ 1 + (1 | tank) + pred + size),
    "b13M01e" = bf(surv | trials(density) ~ 1 + (1 | tank) + pred + size + pred * size))
})

a_file <- here::here("fits", "b13M01list.qs")
models$fits <- qread(a_file)
# create the fits
# models$fits <- purrr::map(.x = models$formulas, .f = function(x) {
#   m <- update(b13M01a,
#               newdata = dataFrogs,
#               formula = x,
#               prior = prior(normal(0, 0.5), class = b))
#   m <- add_criterion(m, c("loo", "waic"))
# })
# qsave(models$fits, a_file)

models$summ <- purrr::map_dfr(.x = models$fits, .f = function(x) {
  spread_draws(model = x, sd_tank__Intercept) %>%
    summarize_draws()
  }, 
  .id = "fit")
models$summ <- models$summ %>%
  mutate(caption = case_when(fit == "b13M01b" ~ "pred",
                   fit == "b13M01c" ~ "size",
                   fit == "b13M01d" ~ "pred + size",
                   fit == "b13M01e" ~ "pred + size + pred:size",
                   TRUE ~ as.character(fit)))
# glimpse(models$summ)

ggplot(models$summ, aes(x = mean, xmin = q5, xmax = q95, y = caption, color = fit)) +
  geom_pointinterval(fatten_point = 3, size = 5) +
  ggrepel::geom_text_repel(aes(label = round(mean, 2)), color= "midnightblue") +
  scale_color_paletteer_d("awtools::bpalette") +
  theme(legend.position = "none") +
  labs(title = "90% interval of sd_tank__Intercept",
       x = "sd_tank__Intercept", y = NULL)
```

The question concerns *the inferred variation variation across tanks*. This means
comparing $\sigma_{tank}$, i.e. $sd_tank__Intercept$, across the models.

If predictors would ocontain all the information then $\sigma_{tank}$ would be
zero.  The results above show that $size$ is a predictor that has little effect
on the outcome.


## 13M2 {-#prac13M2}

```{r}
loo_compare(models$fits$b13M01b, models$fits$b13M01c, models$fits$b13M01d, 
            models$fits$b13M01e, criterion = "waic")
```

b13M01c is the model with $size$ only which confirms our conclusion in 13M1
that it is the least informative model.


Note however, that other predictors in combination with $size$ might actually
make it more relevant. So the above is not an absolute conclusion.


## 13M3 {-#prac13M3}


