
# `logit()` distribution {-#app-logit}

McElrath and Kurz create `logistic()`, `logit()`, etc. Actually
both function exist in the base R package `stats` under the name
`plogis()` for `logistic()` and `qlogis()` for `logit()`.

See [Roman Cheplyaka](https://ro-che.info/articles/2018-08-11-logit-logistic-r)
for a nice explanation.



## `logistic()` / `inv_logit()` function

> This function is in the base r package `stats` and is called `plogis()`.

In McElrath's [GitHub page](https://github.com/rmcelreath/rethinking/blob/a309712d904d1db7af1e08a76c521ab994006fd5/R/distributions.r), there are 2 functions, `logistic()` and `inv_logit()` who are identical
and whose use (inverting the logit function) is the same.

What is the reason?  Shouldn't there be only one function to avoid errors and simplify
code maintenance?

In the Kurz's [bookdown](https://bookdown.org/content/3890/monsters-and-mixtures.html#adding-predictor-variables.)
the same function is called `logistic()`. and Kurz mentions he couldn't find an equivalent function
in `brms`.


As a check, we will use the result of the model 11.1 which is called
`m11.1` by McElrath with `rethinking` and b11.1 by Kurz with `brms`.

```{r}
intercepts <- c(-1.92, -1.27, -0.72, 0.25, 0.89, 1.77)
```


the logistic function as programmed by McElrath and Kurz is

```{r}
logistic <- function(x) {
  p <- 1 / (1 + exp(-x))
  p <- ifelse(x == Inf, 1, p)
  p
}
```

and the `stats::plogis()` gives the same result

```{r}
x <- stats::plogis(intercepts)
y <- logistic(intercepts)
x
stopifnot(identical(x, y))
```


## `logit()` function

> This function is in the base r package `stats` and is called `qlogis()`.

The logit function is the quantile function for the logistic distribution.
Therefore we can use `qlogis` which will return the original intercepts.


```{r}
intercepts <- c(-1.92, -1.27, -0.72, 0.25, 0.89, 1.77)
x <- logistic(intercepts)
x.inv <- stats::qlogis(x)
x.inv
stopifnot(all.equal(x.inv, intercepts))
```


## `dordlogit()` function

The `dordlogit()` is programmed by McElrath's [GitHub page](https://github.com/rmcelreath/rethinking/blob/a309712d904d1db7af1e08a76c521ab994006fd5/R/distributions.r)
as follows

```{r}
dordlogit <- function( x , phi , a , log=FALSE ) {
    a <- c( as.numeric(a) , Inf )
    p <- logistic( a[x] - phi )
    na <- c( -Inf , a )
    np <- logistic( na[x] - phi )
    p <- p - np
    if ( log==TRUE ) p <- log(p)
    p
}
```



Using the `stats::plogis()`, it can be re-written in a more compact form as

```{r}
dordlogit.new <- function(x, phi = 0L, log = FALSE) {
  x <- sort(x)  # the ordering is important
  p <- stats::plogis(q = c(x, Inf), location = phi)
  p <- c( p[1], p[2:length(p)] - p[1:(length(p)-1)] )
  if (log) p <- log(p)
  p
}
```

which can be tested, again using the results of m11.1 (McElrath), b11.1 (Kurz)

```{r}
intercepts <- c(-1.92, -1.27, -0.72, 0.25, 0.89, 1.77)

pk <- dordlogit(1:7, phi = 0, a = intercepts)
pk

pk.new <- dordlogit.new(x = intercepts)
pk.new

stopifnot(identical(pk, pk.new))
```



## `pordlogit()` function

Also `pordlogit()` is given by McElrath as

```{r}
pordlogit <- function( x , phi , a , log=FALSE ) {
    a <- c( as.numeric(a) , Inf )
    if ( length(phi) == 1 ) {
        p <- logistic( a[x] - phi )
    } else {
        p <- matrix( NA , ncol=length(x) , nrow=length(phi) )
        for ( i in 1:length(phi) ) {
            p[i,] <- logistic( a[x] - phi[i] )
        }
    }
    if ( log==TRUE ) p <- log(p)
    p
}
```


which can be rewritten in a much simpler, and probably faster, form as

```{r}
pordlogit.new <- function(x, phi = 0L, log = FALSE) {
  p <- t(
      sapply(X = phi, 
             FUN = function(y) stats::plogis(q = x, location = y, log.p = log)
             )
      )
  if (length(phi) == 1) p <- as.vector(p)
  p
}
```

and we test it as follows

```{r}
# test values when several phi and x
phi <- 1:3
x <- c(0.15, 0.30, 0.50, 0.70, 0.85)

# using McElrath version
m <- pordlogit(x = 1:5, phi = phi, a = x, log = FALSE)
m

# using the new, simpler form
m.new <- pordlogit.new(x, phi, log = FALSE)
m.new

# both should give the same thing
stopifnot(identical(m, m.new))


# and test with only 1 phi value
# using McElrath version
m1 <- pordlogit(x = 1:5, phi = 1, a = x, log = FALSE)
m1

# using the new form
m1.new <- pordlogit.new(x, phi = 1, log = FALSE)
m1.new

stopifnot(identical(m1, m1.new))


# and testing with log == TRUE
# using McElrath version
m.log <- pordlogit(x = 1:5, phi = phi, a = x, log = TRUE)
m.log

# using the new form
m.log.new <- pordlogit.new(x, phi, log = TRUE)
m.log.new

stopifnot(all.equal(m.log, m.log.new))
```


## `rordlogit()` function


Using the results from above we can also simplify `rordlogit()`.

```{r}
rordlogit <- function( n , phi=0 , a) {
    a <- c( as.numeric(a) , Inf )
    k <- 1:length(a)
    
    if ( length(phi)==1 ) {
        p <- dordlogit( k , a=a , phi=phi , log=FALSE )
        y <- sample( k , size=n , replace=TRUE , prob=p )
    } else {
        # vectorized input
        y <- rep(NA,n)
        if ( n > length(phi) ) {
            # need to repeat some phi values
            phi <- rep(phi, ceiling(n/length(phi)) )
        }
        for ( i in 1:n ) {
            p <- dordlogit( k , a=a , phi=phi[i] , log=FALSE )
            y[i] <- sample( k , size=1 , replace=TRUE , prob=p )
        }
    }
    y
}
```


The simpler, and probably faster, format is as follows

```{r}
rordlogit.new <- function(n, x, phi = 0L) {
  
  k <- seq_len(length(x) + 1)
  
  # need to repeat some phi values
  if (n > length(phi)) phi <- rep(phi, ceiling(n/length(phi)))
  
  sapply(X = seq_len(n),
         FUN = function(i) {
           p <- dordlogit.new(x = x, phi = phi[i], log = FALSE)
           sample(k, size = 1, replace = TRUE, prob = p)
           })
}
```


and we test it

```{r}
intercepts <- c(-1.92, -1.27, -0.72, 0.25, 0.89, 1.77)

set.seed(1)
s <- rordlogit(n = 5, phi = 0:3, a = intercepts)
s

set.seed(1)
s.new <- rordlogit.new(n = 5, x = intercepts, phi = 0:3)
s.new

stopifnot(identical(s, s.new))


# and testing it with only 1 phi and n = 1
set.seed(1)
s1 <- rordlogit(n = 1, phi = pi, a = intercepts)
s1

set.seed(1)
s1.new <- rordlogit.new(n = 1, x = intercepts, phi = pi)
s1.new

stopifnot(identical(s1, s1.new))
```


