```{r include=FALSE}
library(rethinking)
library(brms)
library(INLA)
library(dplyr, quietly = TRUE)
library(tidyr, quietly = TRUE)
library(tidybayes, quietly = TRUE)
library(ggdist, quietly = TRUE)
```


# Math notes {-#app-maths}

This appendix is covers a sundry list of mathematical questions, issues and
notations.


## Mode {-#app-maths-mode}

Technically, the mode is defined as the most likely value.  Usually
the following function does precisely that.


```{r}
# find the mode
mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
  }
```


However `ggdist::Mode()` which is used in all `ggdist::mode_xxxx()` functions
gives a different number which is, in the case of a continuous variable, the
value with the highest probability density.

```{r}
data(Howell1)
d <- Howell1 %>%
  filter(age  >= 18)
rm(Howell1)
the_modes <- c("mode" = mode(d$height), "ggdist" = ggdist::Mode(d$height)) 
ggplot(d, aes(x = height)) +
  geom_density() +
  geom_vline(xintercept = the_modes, color = c("red", "blue")) +
  geom_text(data = data.frame(x = the_modes, y = c(0.01, 0.02)),
            aes(x = x, y = y, label = x), color = c("red", "blue"), 
            inherit.aes = FALSE) +
  theme_light() +
  theme(panel.grid = element_blank()) +
  labs(title = "Compare mode() and ggdist::Mode()",
       subtitle = sprintf("ggdist::Mode = %.6f\nmode = %f", 
                          the_modes["ggdist"], the_modes["mode"]))
```

As shown just above, when plotting the density, `ggdist::Mode()` always gives the
right position with maximum density.  The custom `mode()` function, as recommended
for example at [mode](https://stackoverflow.com/questions/2547402/how-to-find-the-statistical-mode)
is pretty much always off.  Often by a lot.

The `ggdist::Mode()` is the function used in this project.


## $\Gamma$ distribution {-#app-gammaexp}

See [wkipedia](https://en.wikipedia.org/wiki/Gamma_distribution) for more details
on the gamma distribution

Gamma distribution is usually expressed in 2 ways, as 
$\Gamma(shape = \alpha, rate = \beta)$ or $\Gamma(shape = \alpha, scale = \theta)$
where $\theta = \frac{1}{\beta}$ and the chosen notation in `INLA` is 
$\Gamma(shape = \alpha, rate = \beta)$.

The exponential and gamma distributions are related in that

$$
\mathcal{Gamma}(1, \beta) = \mathcal{Exponential}(\beta)
$$

