```{r include=FALSE}
library(rethinking)
library(INLA)
library(brms)
library(eflINLA)
library(dplyr, quietly = TRUE)
library(tidyr, quietly = TRUE)
library(modelr, quietly = TRUE)
library(tidybayes, quietly = TRUE)
library(ggdist, quietly = TRUE)
library(paletteer, quietly = TRUE)
library(patchwork, quietly = TRUE)
```



# `eflINLA` {-#eflINLA}

This appendix explains the `eflINLA` package.  It is the vignette project
for that package.

The main use `eflINLA` is to obtain the posterior samples that are converted
to `draws` and `rvar` objects as implemented by the `posterior` package. Once
this is done, the entire toolkit offered by `posterior` greatly simplifies the
data wrangling chores.

We will use the data from the `Howell1` data set from the `rethinking` package.
The way it is used here is largely inspired from Practice 4M7 of chapter 4 from
@elreath2020.

```{r}
data(Howell1)
data04M07 <- Howell1 %>%
  filter(age  >= 18) %>%
  mutate(weight_c = as.vector(scale(weight, center = TRUE, scale = FALSE)))
rm(Howell1)
# skimr::skim(data04M07)
```

## Posterior draws {#eflINLA-posterior}

There are at least 2 ways to extract the posterior distributions from an `inla`
object:
1. From the marginal distribution found in `marginals.fixed` and `marginals.hyperpar`
of an `inla` object or
2. using the function `INLA::inla.posterior.sample`.

To distinguish between the 2, `eflINLA` has 2 functions

1. `draw_marginals`: Draw samples using the `marginals.*` of an `inla` object.
2. `draw_posterior`: Draw samples using `INLA::INLA::inla.posterior.sample()`.

Both of these functions return a `draw_rvars` object which can then be used 
with the `posterior package`.

Note that, technically, a marginal distribution such as `marginals.fixed` is *not* a posterior
distribution.  However, in the case of `INLA` the 2 are extremely close. This is
possible by the `INLA` assumption that the model is part of the class of 
*latent Gaussian Markov random field* (section 1.6, p. 4 of @gomez2020).

