```{r include=FALSE}
library(rethinking)
library(INLA)
library(brms)
library(eflINLA)
library(dplyr, quietly = TRUE)
library(tidyr, quietly = TRUE)
library(modelr, quietly = TRUE)
library(tidybayes, quietly = TRUE)
library(ggdist, quietly = TRUE)
library(paletteer, quietly = TRUE)
library(patchwork, quietly = TRUE)
```


# `eflINLA` {-#eflINLA}

This appendix explains the `eflINLA` package.  It is the vignette project
for that package.

The main use `eflINLA` is to obtain the posterior samples that are converted
to `draws` and `rvar` objects as implemented by the `posterior` package. Once
this is done, the entire toolkit offered by `posterior` greatly simplifies the
data wrangling chores.

We will use the data from the `Howell1` data set from the `rethinking` package.
The way it is used here is largely inspired from Practice 4M7 of chapter 4 from
@elreath2020.

```{r}
data(Howell1)
data04M07 <- Howell1 %>%
  filter(age  >= 18) %>%
  mutate(weight_c = as.vector(scale(weight, center = TRUE, scale = FALSE)))
rm(Howell1)
# skimr::skim(data04M07)
```

```{r}
# the model from practice 4M7 using quap
a_file <- here::here("fits", "m04M07ctr.rds")
m04M07ctr <- readRDS(file = a_file)
# the model from practice 4M7 using inla
a_file <- here::here("fits", "i04M07ctr.rds")
i04M07ctr <- readRDS(file = a_file)
```



## Introduction {-#eflINLA-intro}

`eflINLA` has the following functions which follows the `tidybayes` paradigm
as close as possible. There is 2 groups of functions as a result:

* Marginals: Use the *marginals.x* properties of the `inla` object, e.g.
`x$marginals.fixed`. They give the **marginal** distributions.
* Posterior: Extract the relevant data from the list of posterior samples
obtained by the `INLA::inla.posterior.sample()` function of `INLA`.

```{r echo=FALSE}
data.frame(
  "type" = c("samples", "linear predictor (fitted)", "expected predictor", "prediction"),
  "marginal" = c("tidy_marg_draws_inla", "linpred_marg_draws_inla", "epred_marg_draws_inla", "predict_marg_draws_inla"),
  "posterior" = c("tidy_draws_inla", "linpred_draws_inla", "epred_draws_inla", "predict_draws_inla")
)
```


All of these functions return a `draw_rvars` object which can then be used 
with the wonderful posterior package, see @R-posterior.

Note that, technically, a marginal distribution such as `marginals.fixed` is *not* a posterior
distribution.  However, in the case of `INLA` the 2 are extremely close. This is
possible by the `INLA` assumption that the model is part of the class of 
*latent Gaussian Markov random field* (section 1.6, p. 4 of @gomez2020).

See also section 2.7 of @gomez2020 on an example using the posterior sample.

Also, see chapter 4 of @moraga2019. In particular, the section 4.4.3 shows 
how to use the different marginals. It indicates that, with `INLA`, 

1. *marginal.linear.predictor* represents the posterior marginals of the 
linear predictors.
2. *marginals.fitted.values* represents the *fitted
values obtained by transforming the linear predictors by the inverse of the
link function*.

The terms *fitted* can be really confusing, hence the use of the `tidyverse`
naming convention which addresses this confusion problem.


## Marginal draws {-#eflINLA-marg}

### Sample posterior: `tidy_marg_draws_inla`

We get can a sample using the marginal probabilities

```{r}
samples <- tidy_marg_draws_inla(i04M07ctr, n = 100)
# samples
```

### Linear predictors: `linpred_marg_draws_inla` {-#eflINLA-marg-linpred}

We predict linear predictors by creating a new data. With `INLA` the original
data will be included with the new data set to `NA`.  A function called
`augment_inla` in the `eflINLA` has been created to simplify the result.  See
more details in the documentation of `eflINLA::augment_inla`.

```{r}
newdata <- data.frame(
  height = NA_real_,
  weight_c = modelr::seq_range(i04M07ctr$.args$data$weight_c, n = 10))
i04M07ctr_aug <- eflINLA::augment_inla(i04M07ctr, newdata = newdata)
# str(i04M07ctr_aug)
```

which allow us to predict with a syntx that is very close to `tidyvayes`

```{r}
preds <- linpred_marg_draws_inla(i04M07ctr_aug$inla, i04M07ctr_aug$new_pos,
                                 n = 100)
# preds
```

If one wishes to use the existing data without actual newdata, then the original
`inla` object can be used as follows

```{r}
preds <- linpred_marg_draws_inla(i04M07ctr, n = 50)
# preds
```


### Fitted predictors: `epred_marg_draws_inla` {-#eflINLA-marg-epred}

For fitted predictors, the exact same process is used as for linear predictors
just above.


```{r}
preds <- epred_marg_draws_inla(i04M07ctr_aug$inla, i04M07ctr_aug$new_pos,
                                 n = 100)
# preds
```

### Posterior predictions: `predict_inla` {-#eflINLA-marg-predict}

The posterior predictions can be obtained using the `INLA::inla.posterior.sample()`
function as described in section 2.7 of @gomez2020. However this will only
gives linear predictors. Therefore we must extract the standard deviation
for each fitted predictor and simulate using the standard deviation for each
predictor.

See the discussion from [Faraway](http://julianfaraway.github.io/brinla/examples/chicago.html)
on how the predictions can be obtained using `INLA`. Note that Faraway mentions
the problem of having density for value less than zero.  This is not the case here
as we used `f(weight_c, model = "clinear", range = c(0, Inf), prior = "logtnormal", param = c(0, 1))`.
See `model = "clinear"` in section 3.2.2 of @gomez2020.

This process is done by the `eflINLA::predict_inla` function.

See section 3.3 in @wang2018 for another discussion on prediction.  The 
discussion is not as informative as Faraway above.

> **Warnings**: This is a work in progress. I think the predictors from
`INLA::inla.posterior.sample` are linear, we should use fitted predictors.
Also, please obtain better references.

```{r}
preds <- predict_inla(i04M07ctr, n = 100)
# preds
```


