--- 
title: "Statistical Rethinking - Solutions"
author: "frank"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [books.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "Packages and setup to use this book."
---


```{r setup, include=FALSE}
library(rethinking)
```


# Introduction {-}

## Session info {-}

```{r}
sessioninfo::session_info()
```


## Prerequisites {-}

To install the `rstan` package the rstan site provides useful guidance at
[rstan site](https://github.com/stan-dev/rstan/wiki/Installing-RStan-on-Windows).

The sequence to follow is

1. Ensure `Rtools` is installed
2. Ensure `rstan` is installed
3. Install rethinking only once

### Install `Rtools` {-}

Make sure `Rtools` is installed before proceeding further.

Get the environment path

```{r}
sys_path <- Sys.getenv("PATH")
sys_path <- unlist(stringr::str_split(string = sys_path, pattern = ";"))

```

then find position of RTools and test if it is installed

```{r}
tmp <- grepl(pattern = "(Rtools)|(RBuildTools)", sys_path, ignore.case = TRUE)
stopifnot(any(tmp))
```

and verify that `g++` can really be called from `R`

```{r}
system("g++ -v")
system("where make")
```


### Install `rstan` {-}

Install rstan and its tools.

```{r eval=FALSE}
install rstan and its tools
install.packages(c("rstan", "rstantools"))
```


and verify that tool chain works

```{r}
fx <- inline::cxxfunction(signature(x = "integer", y = "numeric" ) ,
                           'return ScalarReal( INTEGER(x)[0] * REAL(y)[0] ) ;' )
```


and just a  final verification to ensure all is correct

```{r}
# this should return 10
fx <- inline::cxxfunction(signature(x = "integer", y = "numeric" ) ,
                           'return ScalarReal( INTEGER(x)[0] * REAL(y)[0] ) ;' )
fx( 2L, 5 )
if(fx( 2L, 5 ) == 10L) {
    msg <- paste("RStan installed with success. Version", packageVersion("rstan"))
    message(msg)
} else {
    stop("RStan installation failed.")
}
```


### Install `rethinking` {-}


```{r eval=FALSE}
install.packages(c("coda", "mvtnorm", "devtools"))
library(devtools)
devtools::install_github("rmcelreath/rethinking")
```


> **Important** `rethinking` mus be reintalled to work under `R 4.0`. 
  But it is not on 'CRAN' and must be reinstalled locally. Go to 
  [solution](see https://github.com/rmcelreath/rethinking/issues/250) to solve it.
  This will show the command on the next lines.

Use the following to install `rethinking` locally when usi `R` with version >= 4.0.

```{r eval=FALSE}
remotes::install_github("rmcelreath/rethinking")
```

## References {-}

These notes borrow heavily from @elreath2016 and  @kurtz2020a which does all the
bayesian analysis with `brms` instead of `rethinking`.

The newest version of `rethinking` is in @elreath2020 and covered by @kurtz2020b.